Towards Detecting Stalkerware Through Network Traces

### 2019-03-20 paper/main.tex
changed quotes

As previously mentioned, spyware has a high turnover rate on the Google Play 
store, which leads to a practice we refer to as “reskinning:” a single app will 
be posted under many names, icons, and manufacturer names. Presumably, 
manufacturers do this because spyware gets reported frequently and taken down - 
we have already seen one instance of a spyware app getting removed from the 
Play store while we were examining it. Overt stalkerware which is not available 
through Google Play has also turned out to use reskinning, for reasons we 
haven’t yet discovered. Apps can frequently be identified as copies of each 
other by observing their dashboards (where users view the communications and 
locations of the target phones), download sites, and advertisements. Several 
groups of off-store apps have nearly identical websites, in terms of 
organization, color, fonts, placement of buttons, prices, and other factors. 
For example, the “ILFMobile group” consists of three different companies 
(ILFMobile, PowerLine Group, and PhoneSpector) that claim to make a total of 
four different apps. Further investigation showed that all three companies had 
the same physical address in addition to identical websites. Unfortunately, 
they also had identical bugs in their payment methods, so we were unable to 
purchase the apps and verify that their network traces were identical. Other 
app sites appear to be similar to each other in other ways, 
with many having a large panel with a darkened image background extolling the 
features of the app above two buttons, one for a "live demo" and the other to 
"buy now." This is somewhat circumstantial evidence - it is entirely possible 
that these manufacturers simply chose similar web templates, or copied each 
other's site in an attempt to undercut competition. We therefore confirmed our 
suspicions by checking if the traffic signatures were also similar. These 
reskinned programs are potentially a boon to 
our investigation as the identical network signatures make it more likely that 
one signature generalizes across multiple apps. We have identified the 
following groups of applications that appear to be reskinned versions of each 
other.
\begin{enumerate}
	\item \textbf{SpyToMobile group}: Employee Work Spy by Piter Cline, SMS 
	Tracker by TheHar, Mobile Tracking by Antwat, Cell Phone Tracker by Susdel, 
	and Texting, Chat, 
	Phone Spy by MaxLo. All of these apps download the same program, which is 
	called SpyToMobile. Any app from this group installed after the first 
	states ``SpyToMobile is already installed.''
	\item \textbf{ILFMobile group}: HighsterMobile, PhoneSpector, Auto Forward 
	Spy, Surepoint Spy, and EasySpy. We were unable to purchase these and 
	ensure that they have the same signatures due to a bug in their payment 
	system that prevented us from purchasing them.
	\item \textbf{Codero group}: HelloSpy and TheTruthSpy have identical 
	network signatures as well as nearly identical websites.
	Spyzie group: Spyzie and SpyMyFone also have identical network signatures 
	as well as identical websites.
	\item \textbf{CallSMSTracker group}: Call and Message Tracker by Remote 
	Karath and Cell Tracker by Trackme. These apps have identical dashboards 
	and contact the same domains.
\end{enumerate}
We cannot guarantee that we have identified all of the members of any 
particular group. If we have not, it will make other members easier to 
identify, since signatures we have already found will scale to identify more 
apps than we had time to download and manually examine.

@@

App manufacturers appear to be well aware of their competition. When the term 
``hello spy'' is typed into Google Search, ads appear captioned ``Hellospy | 
See All Phone Data Remotely | mSpy.com'' and ``HelloSpy - Free Download | \#1 
Phone Tracker Software\&App | spymyfone.com.'' Clicking on these leads the user 
to the download pages for mSpy and SpyMyFone respectively, not the site for 
HelloSpy. Based on network traces and website appearances, we do not believe 
any of these three apps are made by the same manufacturer. Spyware groups also 
create “review” websites that review their own products, posing as third party 
spyware enthusiasts. At least one blog that is supposedly run by 
an independent party reviewed apps that coincidentally were all made by the 
ILFMobile group\cite{noauthor_best_nodate}. This blog reviews the 
“top 5 cell phone spy apps,” all of which are ILFMobile’s own apps, on a 
website called “bestcellphonespyapps.com.” We haven’t determined what advantage 
spyware manufacturers derive from appearing to make more apps and competing 
with themselves.

@@

\section{Related Work}
Previous studies [1,2], the media [3,4], and intimate partner violence (IPV) 
survivors have all reported that spyware is becoming an increasing source of 
intimate partner surveillance (IPS). Tracking functions supported by spyware 
allow stalkers to monitor communications, track locations, and remotely 
activate cameras and microphones [3,4]. In Dr. Ristenpart’s work [5], they 
proposed methods for searching IPS-related apps. By Google search query 
expansion and unsupervised classification, they collected 9,224 apps possibly 
related to IPS on Google play and open-web. They manually selected 70 apps for 
investigation. They observed the most fundamental functions of these apps are 
monitoring locations, communication logs and data, media contents, and phone 
usages. They also found that existing anti-spyware tools cannot effectively 
detect IPS-related apps.
Mobile apps are often identified via the User-Agent string of the HTTP request, 
when one is available [7]. With the increased use of HTTPS this may not always 
be the case, but we expect to find (based on conversations with authors of 
previous work) that security and best coding practices are not prioritized by 
stalkerware developers. Malware identification may be of limited use as well, 
since techniques are available to cluster malware that comes from the same 
source code into “families.” [8]

@@

\begin{table*}
	\begin{tabular}{p{5cm}p{5cm}p{5cm}}
		\toprule
		Generalized Signature & Unique Apps & Total Apps \\
		\hline
		Domain contains “spy,” “track,” “find” & 6 & 12 \\
		SNI contains “spy,” “track,” “find” & 6 & 6 \\
		User-Agent contains “spy,” “track,” “find” & 1 & 3 \\
		User-Agent contains known app name & 1 & 1 \\
		Domain contains known app name & 3 & 4 \\
		SNI contains known app name & 1 & 1 \\
		SNI contains known app manufacturer & 3 & 3 \\
		Domain contains known app manufacturer & 3 & 3 \\
		\midrule
	\end{tabular}
\end{table*}

In the subset of apps we studied, the words “spy,” “find,” and “track” appeared 
frequently in various fields of the traffic. The most common place to find them 
was in the destination domain, which we found using Wireshark’s reverse DNS 
lookup tool. 12 of the 26 apps that we eventually found signatures for 
contained these spyware keywords in their domains. 
For the apps whose traffic doesn’t contain “spy,” “find,” or “track,” we will 
next check for the app name and manufacturer name or substrings of these two. 
Using these indicators has a significant disadvantage - we must be certain that 
the app is stalkerware and that the manufacturer doesn’t make anything except 
stalkerware. In the case of smaller apps and reskinned apps, this may actually 
be a reasonable assumption - manufacturer names are generated to fool the 
Google Play store into thinking the app is new, not to identify a real company, 
and are therefore unique by necessity. We note that the two apps we identify 
only by manufacturer are made by GPSWOX and GeoZilla. GPSWOX makes 
location-tracking apps of all kinds, some of which may be difficult to use as 
stalkerware, but GeoZilla’s only apps are all designed to track a phone. Since 
only two apps were identified this way, the false positive rate should be 
investigated further.
We identified the following generalized signatures that can be used to identify 
most of our apps given nothing but information available on the Google Play 
store. The second step of our analysis will be to search for these general 
signatures in our personal phones’ traffic captures. This will give a better 
idea of whether it is possible to use general signatures without generating a 
prohibitive number of false positive stalkerware detections.

@@

The majority of stalkerware apps communicate though HTTP or HTTPS, while some 
use their own protocols. We examined the domain, SNI, User-Agent, and HTTP 
host, for each app. The words “spy,” “find,” and “track” appeared frequently in 
various fields of the traffic. The most common place to find them was in the 
destination domain, which we found using Wireshark’s reverse DNS lookup tool. 
12 of the 26 apps that we eventually found signatures for contained these 
spyware keywords in their domains. We identified the following generalized 
signatures that can be used to identify most of our apps given nothing but 
information available on the Google Play store.

@@

Previously, we had expressed an interest in seeing if the signatures we found 
by hand could identify more apps without manual inspection. There are (as far 
as we can see) two possible methods of doing this. The first is to 
auto-generate signatures based on a list of keywords that commonly appear in 
domains, SNIs, User-Agents, and HTTP host names for stalkerware traffic. These 
keywords might include the names of the apps, substrings that appear in the app 
names, and words that seem relevant to tracking apps. A more complicated method 
would be to set up bots that can download potential spyware by searching 
various terms on the app store like “tracker app.” The bots would have to run 
the apps and monitor their traffic, identifying app traffic by eliminating 
traffic from other sources. The remaining traffic would belong to the app, and 
the domains, SNIs, user-agents, and any other available identifying features 
could be recorded. However, the most difficult problem in this case becomes 
determining which apps are actually stalkerware on the app store, which is 
beyond the scope of this project.

>>>>

As previously mentioned, spyware has a high turnover rate on the Google Play 
store, which leads to a practice we refer to as "reskinning:" a single app will 
be posted under many names, icons, and manufacturer names. Presumably, 
manufacturers do this because spyware gets reported frequently and taken down - 
we have already seen one instance of a spyware app getting removed from the 
Play store while we were examining it. Overt stalkerware which is not available 
through Google Play has also turned out to use reskinning, for reasons we 
haven't yet discovered. Apps can frequently be identified as copies of each 
other by observing their dashboards (where users view the communications and 
locations of the target phones), download sites, and advertisements. Several 
groups of off-store apps have nearly identical websites, in terms of 
organization, color, fonts, placement of buttons, prices, and other factors. 
For example, the "ILFMobile group" consists of three different companies 
(ILFMobile, PowerLine Group, and PhoneSpector) that claim to make a total of 
four different apps. Further investigation showed that all three companies had 
the same physical address in addition to identical websites. Unfortunately, 
they also had identical bugs in their payment methods, so we were unable to 
purchase the apps and verify that their network traces were identical. Other 
app sites appear to be similar to each other in other ways, 
with many having a large panel with a darkened image background extolling the 
features of the app above two buttons, one for a "live demo" and the other to 
"buy now." This is somewhat circumstantial evidence - it is entirely possible 
that these manufacturers simply chose similar web templates, or copied each 
other's site in an attempt to undercut competition. We therefore confirmed our 
suspicions by checking if the traffic signatures were also similar. These 
reskinned programs are potentially a boon to 
our investigation as the identical network signatures make it more likely that 
one signature generalizes across multiple apps. We have identified the 
following groups of applications that appear to be reskinned versions of each 
other.
\begin{enumerate}
	\item \textbf{SpyToMobile group}: Employee Work Spy by Piter Cline, SMS 
	Tracker by TheHar, Mobile Tracking by Antwat, Cell Phone Tracker by Susdel, 
	and Texting, Chat, 
	Phone Spy by MaxLo. All of these apps download the same program, which is 
	called SpyToMobile. Any app from this group installed after the first 
	states ``SpyToMobile is already installed.''
	\item \textbf{ILFMobile group}: HighsterMobile, PhoneSpector, Auto Forward 
	Spy, Surepoint Spy, and EasySpy. We were unable to purchase these and 
	ensure that they have the same signatures due to a bug in their payment 
	system that prevented us from purchasing them.
	\item \textbf{Codero group}: HelloSpy and TheTruthSpy have identical 
	network signatures as well as nearly identical websites.
	Spyzie group: Spyzie and SpyMyFone also have identical network signatures 
	as well as identical websites.
	\item \textbf{CallSMSTracker group}: Call and Message Tracker by Remote 
	Karath and Cell Tracker by Trackme. These apps have identical dashboards 
	and contact the same domains.
\end{enumerate}
We cannot guarantee that we have identified all of the members of any 
particular group. If we have not, it will make other members easier to 
identify, since signatures we have already found will scale to identify more 
apps than we had time to download and manually examine.

@@

App manufacturers appear to be well aware of their competition. When the term 
``hello spy'' is typed into Google Search, ads appear captioned ``Hellospy | 
See All Phone Data Remotely | mSpy.com'' and ``HelloSpy - Free Download | \#1 
Phone Tracker Software\&App | spymyfone.com.'' Clicking on these leads the user 
to the download pages for mSpy and SpyMyFone respectively, not the site for 
HelloSpy. Based on network traces and website appearances, we do not believe 
any of these three apps are made by the same manufacturer. Spyware groups also 
create "review" websites that review their own products, posing as third party 
spyware enthusiasts. At least one blog that is supposedly run by 
an independent party reviewed apps that coincidentally were all made by the 
ILFMobile group\cite{noauthor_best_nodate}. This blog reviews the 
"top 5 cell phone spy apps," all of which are ILFMobile's own apps, on a 
website called "bestcellphonespyapps.com." We haven't determined what advantage 
spyware manufacturers derive from appearing to make more apps and competing 
with themselves.

@@

\section{Related Work}
Previous studies [1,2], the media [3,4], and intimate partner violence (IPV) 
survivors have all reported that spyware is becoming an increasing source of 
intimate partner surveillance (IPS). Tracking functions supported by spyware 
allow stalkers to monitor communications, track locations, and remotely 
activate cameras and microphones [3,4]. In Dr. Ristenpart's work [5], they 
proposed methods for searching IPS-related apps. By Google search query 
expansion and unsupervised classification, they collected 9,224 apps possibly 
related to IPS on Google play and open-web. They manually selected 70 apps for 
investigation. They observed the most fundamental functions of these apps are 
monitoring locations, communication logs and data, media contents, and phone 
usages. They also found that existing anti-spyware tools cannot effectively 
detect IPS-related apps.
Mobile apps are often identified via the User-Agent string of the HTTP request, 
when one is available [7]. With the increased use of HTTPS this may not always 
be the case, but we expect to find (based on conversations with authors of 
previous work) that security and best coding practices are not prioritized by 
stalkerware developers. Malware identification may be of limited use as well, 
since techniques are available to cluster malware that comes from the same 
source code into "families." [8]

@@

\begin{table*}
	\begin{tabular}{p{5cm}p{5cm}p{5cm}}
		\toprule
		Generalized Signature & Unique Apps & Total Apps \\
		\hline
		Domain contains "spy," "track," "find" & 6 & 12 \\
		SNI contains "spy," "track," "find" & 6 & 6 \\
		User-Agent contains "spy," "track," "find" & 1 & 3 \\
		User-Agent contains known app name & 1 & 1 \\
		Domain contains known app name & 3 & 4 \\
		SNI contains known app name & 1 & 1 \\
		SNI contains known app manufacturer & 3 & 3 \\
		Domain contains known app manufacturer & 3 & 3 \\
		\midrule
	\end{tabular}
\end{table*}

In the subset of apps we studied, the words "spy," "find," and "track" appeared 
frequently in various fields of the traffic. The most common place to find them 
was in the destination domain, which we found using Wireshark's reverse DNS 
lookup tool. 12 of the 26 apps that we eventually found signatures for 
contained these spyware keywords in their domains. 
For the apps whose traffic doesn't contain "spy," "find," or "track," we will 
next check for the app name and manufacturer name or substrings of these two. 
Using these indicators has a significant disadvantage - we must be certain that 
the app is stalkerware and that the manufacturer doesn't make anything except 
stalkerware. In the case of smaller apps and reskinned apps, this may actually 
be a reasonable assumption - manufacturer names are generated to fool the 
Google Play store into thinking the app is new, not to identify a real company, 
and are therefore unique by necessity. We note that the two apps we identify 
only by manufacturer are made by GPSWOX and GeoZilla. GPSWOX makes 
location-tracking apps of all kinds, some of which may be difficult to use as 
stalkerware, but GeoZilla's only apps are all designed to track a phone. Since 
only two apps were identified this way, the false positive rate should be 
investigated further.
We identified the following generalized signatures that can be used to identify 
most of our apps given nothing but information available on the Google Play 
store. The second step of our analysis will be to search for these general 
signatures in our personal phones' traffic captures. This will give a better 
idea of whether it is possible to use general signatures without generating a 
prohibitive number of false positive stalkerware detections.

@@

The majority of stalkerware apps communicate though HTTP or HTTPS, while some 
use their own protocols. We examined the domain, SNI, User-Agent, and HTTP 
host, for each app. The words "spy," "find," and "track" appeared frequently in 
various fields of the traffic. The most common place to find them was in the 
destination domain, which we found using Wireshark's reverse DNS lookup tool. 
12 of the 26 apps that we eventually found signatures for contained these 
spyware keywords in their domains. We identified the following generalized 
signatures that can be used to identify most of our apps given nothing but 
information available on the Google Play store.

@@

Previously, we had expressed an interest in seeing if the signatures we found 
by hand could identify more apps without manual inspection. There are (as far 
as we can see) two possible methods of doing this. The first is to 
auto-generate signatures based on a list of keywords that commonly appear in 
domains, SNIs, User-Agents, and HTTP host names for stalkerware traffic. These 
keywords might include the names of the apps, substrings that appear in the app 
names, and words that seem relevant to tracking apps. A more complicated method 
would be to set up bots that can download potential spyware by searching 
various terms on the app store like "tracker app." The bots would have to run 
the apps and monitor their traffic, identifying app traffic by eliminating 
traffic from other sources. The remaining traffic would belong to the app, and 
the domains, SNIs, user-agents, and any other available identifying features 
could be recorded. However, the most difficult problem in this case becomes 
determining which apps are actually stalkerware on the app store, which is 
beyond the scope of this project.
